An Asymmetry-Aware Method for Focal Mechanism and Uncertainty Estimation in Single-seismometer Settings

### 2025-11-11 project/paper/paper.tex
back at it again
Farallon is down

\title{Single-seismometer Focal Mechanism and Uncertainty Estimation from Body Waves}
\shorttitle{Single-seismometer Focal Mechanism and Uncertainty Estimation} % used for header, not mandatory

@@

    \begin{align} \label{eq:2}
      \nonumber s_R &= 0.5 \sin(\lambda) \sin(2\delta)\\
      \nonumber q_R &= \sin(\lambda) \cos(2\delta) \sin(\psi_r) + \cos(\lambda) \cos(\delta) \cos(\psi_r)\\
      \nonumber p_R &= \cos(\lambda) \sin(\delta) \sin(2\psi_r) - 0.5\sin(\lambda) \sin(2\delta) \cos(2\psi_r)\\
      \nonumber p_L &= 0.5\sin(\lambda) \sin(2\delta) \sin(2\psi_r) + \cos(\lambda) \sin(\delta) \cos(2\psi_r)\\
      q_L &= -\cos(\lambda) \cos(\delta) \sin(\psi_r) + \sin(\lambda) \cos(2\delta) \cos(\psi_r)
    \end{align}

@@

        \foreach \p in {2,3}
          \foreach \h in {1,2,3,4,5}
            \draw (P\p) -- (H\h);

@@

        \foreach \a in {1,2,3}
          \draw (A\a) -- (M);

        % \draw[dashed] (P1) -- (H1);
        \draw (P1) -- (H2);
        \draw (P1) -- (H3);
        \draw (P1) -- (H4);
        \draw (P1) -- (H5);

        \draw (H1) -- (A3);
        \draw (H1) -- (A1);
        \draw (H1) -- (A2);
        \draw (H2) -- (A1);
        \draw (H2) -- (A2);
        \draw (H3) -- (A1);
        \draw (H3) -- (A2);
        \draw (H4) -- (A3);
        \draw (H5) -- (A3);
      \end{tikzpicture}
      \caption{Computational graph of forward model (edges are non-zero partial derivatives).}
      \label{fig:1}
    \end{figure}

@@

    i) From $\bm{m}$ to $\bm{\eta}$, we can use equation \ref{eq:2} to define the Jacobian
    as follows:
    \begin{align*}
      &a = \sin(\lambda)\;\;\;\;\;\;\;\;\;\;\;\;\; b = \cos(2\delta)\;\;\;\;\;\;\;\;\;\;\; c = \cos(\lambda)\\
      &d = \sin(2\delta)\;\;\;\;\;\;\;\;\;\;\;\; e = \sin(\delta)\;\;\;\;\;\;\;\;\;\;\;\; f = \cos(\delta)\\
      &g = \sin(\psi - \phi)\;\;\;\;\;\; h = \cos(\psi - \phi)\;\;\;\;\; p = cos(2(\psi - \phi))\\
      &q = \sin(2(\psi - \phi))
    \end{align*}
    \begin{align} \label{eq:18}
      \nonumber&\bm{J_p}(\bm{\eta}) = \\
      &\begin{bmatrix}
        0         &ab         &0.5cd\\
        abh-cfg   &-2adg-ceh  &cbg-afh\\
        2cep+adq  &cfq-abp    &-aeq-0.5cdp\\
        -cfh-abg  &ceg-2adh   &afg+cbh\\
        adp-2ceq  &abq+cfp    &0.5cdq-aep
      \end{bmatrix}
    \end{align}

@@

    Combining \ref{eq:18}, \ref{eq:19} and \ref{eq:20} gives the overall gradient:
    \begin{align} \label{eq:21}
      \nabla_{\bm{m}}\Phi = \bm{J_p}(\bm{\eta})^\top \bm{J_\eta}(\bm{A})^\top \nabla_{\bm{A}}\tilde{\Phi}
    \end{align}

>>>>

% \title{Single-seismometer Focal Mechanism and Uncertainty Estimation from Body Waves}
\title{An Asymmetry-Aware Method for Focal Mechanism and Uncertainty Estimation in Single-seismometer Settings}
\shorttitle{Single-seismometer Focal Mechanism and Uncertainty Estimation} % used for header, not mandatory

@@

    \begin{align} \label{eq:2}
      \nonumber s_R &= 0.5 \sin(\lambda) \sin(2\delta)\\
      \nonumber q_R &= \sin(\lambda) \cos(2\delta) \sin(\psi_r) + \cos(\lambda) \cos(\delta) \cos(\psi_r)\\
      \nonumber p_R &= \cos(\lambda) \sin(\delta) \sin(2\psi_r) - 0.5\sin(\lambda) \sin(2\delta) \cos(2\psi_r)\\
      \nonumber p_L &= 0.5\sin(\lambda) \sin(2\delta) \sin(2\psi_r) + \cos(\lambda) \sin(\delta) \cos(2\psi_r)\\
      \nonumber q_L &= -\cos(\lambda) \cos(\delta) \sin(\psi_r) + \sin(\lambda) \cos(2\delta) \cos(\psi_r)\\
      \psi_r &= \psi - \phi
    \end{align}

@@

        \foreach \p in {2,3}
          \foreach \h in {1,2,3,4,5}
            \draw[->] (P\p) -- (H\h);

@@

        \foreach \a in {1,2,3}
          \draw[->] (A\a) -- (M);

        % \draw[dashed] (P1) -- (H1);
        \draw[->] (P1) -- (H2);
        \draw[->] (P1) -- (H3);
        \draw[->] (P1) -- (H4);
        \draw[->] (P1) -- (H5);

        \draw[->] (H1) -- (A1);
        \draw[->] (H1) -- (A2);
        \draw[->] (H2) -- (A1);
        \draw[->] (H2) -- (A2);
        \draw[->] (H3) -- (A1);
        \draw[->] (H3) -- (A2);
        \draw[->] (H4) -- (A3);
        \draw[->] (H5) -- (A3);
      \end{tikzpicture}
      \caption{Computation graph of forward model (edges are non-zero partial derivatives).}
      \label{fig:1}
    \end{figure}

@@

    i) From $\bm{m}$ to $\bm{\eta}$, we can use equation \ref{eq:2} to define the Jacobian
    as follows:
    \begin{align*}
      &a = \sin(\lambda)\;\;\;\;\;\;\;\;\;\;\;\;\; b = \cos(2\delta)\;\;\;\;\;\;\;\;\;\;\; c = \cos(\lambda)\\
      &d = \sin(2\delta)\;\;\;\;\;\;\;\;\;\;\;\; e = \sin(\delta)\;\;\;\;\;\;\;\;\;\;\;\; f = \cos(\delta)\\
      &g = \sin(\psi_r)\;\;\;\;\;\;\;\;\;\;\; h = \cos(\psi_r)\;\;\;\;\;\;\;\;\;\; p = cos(2(\psi_r)\\
      &q = \sin(2\psi_r)
    \end{align*}
    \begin{align} \label{eq:18}
      \nonumber&\bm{J_m}(\bm{\eta}) = \\
      &\begin{bmatrix}
        0         &ab         &0.5cd\\
        abh-cfg   &-2adg-ceh  &cbg-afh\\
        2cep+adq  &cfq-abp    &-aeq-0.5cdp\\
        -cfh-abg  &ceg-2adh   &afg+cbh\\
        adp-2ceq  &abq+cfp    &0.5cdq-aep
      \end{bmatrix}
    \end{align}

@@

    Combining \ref{eq:18}, \ref{eq:19} and \ref{eq:20} gives the overall gradient:
    \begin{align} \label{eq:21}
      \nabla_{\bm{m}}\Phi = \bm{J_m}(\bm{\eta})^\top \bm{J_\eta}(\bm{A})^\top \nabla_{\bm{A}}\tilde{\Phi}
    \end{align}
